The Way Forward in Parallelizing Dynamic Languages

### 2014-04-29 talk/icooolps2014/position-paper.tex
some changes

\subsection{Why is there a GIL?}
The GIL is a very simple synchronization mechanism for supporting
multi-threading in the interpreter. The basic guarantee is that the
GIL may only be released in-between bytecode instructions. The
interpreter can thus rely on complete isolation and atomicity of these
instructions. As a consequence, applications can rely on certain
operations to be atomic. While this is probably not a good idea,
it is used in practice. A solution replacing the GIL should therefore
uphold these guarantees, while preferably also be as easily
implementable as a GIL for the interpreter.

@@

The list of criterias for evaluating potential solutions is as follows:
\begin{itemize}
\item Performance
\item Changes required to existing applications
\item Better synchronization options for applications (e.g. atomic blocks)
\item Ease of implementation (interpreter-level)
\item Integration with external libraries
\end{itemize}

\subsubsection*{fine-grained locking}

@@

\subsubsection*{Shared-nothing / multiple processes}

@@

\subsubsection*{Transactional Memory}
\paragraph{HTM}

@@

\appendix
\section{Appendix Title}

This is the text of the appendix, if you need one.

>>>>

\subsection{Why is there a GIL?}
The GIL is a very simple synchronization mechanism for supporting
multi-threading in the interpreter. The basic guarantee is that the
GIL may only be released in-between bytecode instructions. The
interpreter can thus rely on complete isolation and atomicity of these
instructions. As a consequence, applications can rely on certain
operations to be atomic. While this is probably not a good idea,
it is used in practice. A solution replacing the GIL should therefore
uphold these guarantees, while preferably also be as easily
implementable as a GIL for the interpreter.
[xxx mention that the interpreter is typically very large and maintained
by open-source communities]

@@

For the discussion we define a set of criteria to evaluate the
multiple potential solutions for removing or avoiding the GIL and its
limitations:

\begin{description}
\item[Performance:] How well does the approach perform compared to the
  GIL on single and multiple threads?
\item[Existing applications:] How big are the changes required to
  integrate with and parallelize existing applications?
\item[Better synchronization:] Does the approach enable better
  synchronization mechanisms for applications (e.g. atomic blocks)?
\item[Implementation:] How difficult is it to implement the approach
  in the interpreter?
\item[External libraries:] Does the approach allow for easy
  integration of external libraries?
\end{description}

\subsubsection{Fine-Grained Locking}

The first obvious candidate to replace the GIL is to use multiple
locks instead of a single global lock. By refining the granularity
of the locking approach, we gain the ability to run code that does
not access the same objects in parallel.

@@

\subsubsection{Shared-Nothing / multiple processes}

@@

\subsubsection{Transactional Memory}
\paragraph{HTM}

@@

%% \appendix
%% \section{Appendix Title}

%% This is the text of the appendix, if you need one.
